% this file is called up by thesis.tex
% content in this file will be fed into the main document

%: ----------------------- name of chapter  -------------------------
\chapter{Conclusions and Recommendations} % top level followed by section, subsection


%: ----------------------- paths to graphics ------------------------

% change according to folder and file names
\ifpdf
    \graphicspath{{5/figures/PNG/}{5/figures/PDF/}{5/figures/}}
\else
    \graphicspath{{5/figures/EPS/}{5/figures/}}
\fi

%: ----------------------- contents from here ------------------------
\section{Conclusion}
In this project, a web application was developed to help travelers find detail information whcih always available on arrival at the various terminals and stations but not readily available to them prior to starting to their journey. To accomplish this project, some terminals and stations were clearly mapped on OpenStreetMap. After mapping, the data was extracted and further processed to create a geodatabase using PostgreSQL and PostGIS. The data was rendered with Leaflet. Search and routing functionality were subsequently added. The developed application has map rendering and text based interaction to input and output features. It can therefore be concluded that the application will improve trip planning and easy access to information only available within terminals to travelers hence saving time and other resources.It could also be adopted by Ghana Tourism Authority to help tourists find their way around Ghana transport network.

The future objective is to deploy the application making it available for public use in Ghana.

\section{Recommendations}
This project recommends that users should be able to book seats from the platform and also support voice input for the visually impaired as well. The system should get users current location and find nearest available departure stations for their routes. Also a commenting system for users to interact and share experiences about each route.

% ---------------------------------------------------------------------------
%: ----------------------- end of thesis sub-document ------------------------
% ---------------------------------------------------------------------------

